---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---

# Get GSOD climate data for Nepal

Historic climate data for available stations in Nepal.

## LOAD LIBRARIES

```{r}
library(knitr)
library(gsod)
library(GSODR)
library(tidyverse)
library(leaflet)
library(lubridate)
library(broom)
#for using splines instead of straight lines in graph
if(file.exists("D:/rwd/geomspline.R")) {source("D:/rwd/geomspline.R")} else
{
  library(RCurl)
  script <- getURL("https://raw.githubusercontent.com/asheshwor/geeshared/master/rcode/geomspline.R", ssl.verifypeer = FALSE)
  eval(parse(text = script))
}
```

## GET LIST OF STATIONS 

```{r eval=FALSE}
load(system.file("extdata", "isd_history.rda", package = "GSODR"))
np <- subset(isd_history, ISO2C == "NP")
kable(np)
```

## GET DATA FROM STATIONS

There are 17 stations listed. Now, fetch data for those stations. *Takes a while to run.*

```{r eval=FALSE}
np.data <- get_GSOD(years = c(1990:2019), station = np$STNID)
head(np.data)
tail(np.data)
nrow(np.data) #37.3k
```

Save data to file.

```{r eval=FALSE}
write.csv(np.data, "np_gsod_1990-2019.csv", row.names = FALSE)
```

Read saved data:

```{r}
np.data <- read.csv("D:/rwd/dataset/np_gsod_1990-2019.csv")
head(np.data)
```

Summary of retrieved data:

```{r}
names(np.data)
data.summary <- np.data %>%
  select(STNID, NAME, YEAR, PRCP) %>% 
  group_by(STNID, NAME, YEAR) %>% 
  tally() %>% 
  arrange(NAME, YEAR)
kable(data.summary)
```

Save output:

```{r}
write.csv(data.summary, "climatedata_2017-2020.csv",
          row.names = FALSE)
```

Check station locations on a map.

```{r}
pu <- paste("<b>STNID:</b>", as.character(np$STNID), "<br>",
            "<b>NAME:</b>", np$NAME, "<br>",
            "<b>DATES:</b>", paste0(np$BEGIN, "-",np$END), "<br>")
#provider = "CartoDB.Positron"
lmap <- leaflet(np) %>% 
  addTiles() %>% 
  addMarkers(popup = pu)
lmap
```

Stations from DHM, Nepal.
http://www.dhm.gov.np/meteorological-station/ accessed 06/08/2020.

```{r}
np.stations <- read.csv("D:/rwd/dataset/dhm_stations.csv")
head(np.stations)
np.stations <- np.stations %>% 
  select(stationname, district, lat, long, elevation,
         agest, nmsest, agclos, nmsclos, remarks) %>% 
  mutate(latitude = as.numeric(substr(lat, 1,2)) +
           as.numeric(substr(lat, 3,4))/60,
         longitude = as.numeric(substr(long, 1, 2)) +
           as.numeric(substr(long, 3, 4))/60) %>% 
  select(-lat, -long)
head(np.stations)
```

Check station locations on a map.

```{r}
pu <- paste("<b>Station name:</b>", as.character(np.stations$stationname), "<br>",
            "<b>District:</b>", np.stations$district, "<br>",
            "<b>Elevation:</b>", np.stations$elevation, "<br>",
            "<b>agest:</b>", np.stations$agest, "<br>",
            "<b>agclos:</b>", np.stations$agclos, "<br>",
            "<b>nmsest:</b>", np.stations$nmsest, "<br>",
            "<b>nmsclos:</b>", np.stations$nmsclos, "<br>",
            "<b>remarks:</b>", np.stations$remarks, "<br>")
#provider = "CartoDB.Positron"
lmap <- leaflet(np.stations) %>% 
  addTiles() %>% 
  addMarkers(popup = pu)
lmap
```

Official monsoon dates for Nepal.

```{r}
dhm.file <- "D:/github/climate-nepal/data/MonsoonOnsetOfficial.csv"
dhm.monsoon <- read.csv(dhm.file, header = TRUE,
                        colClasses = c("numeric", "character",
                                         "character", "numeric",
                                         "numeric", "numeric"),
                        fileEncoding="UTF-8-BOM")
dhm.monsoon <- dhm.monsoon %>% 
  select(year, month, day, monthend, dayend) %>% 
  mutate(monsoon = as.Date(paste(year, month, day,
                                 sep = "-"), "%Y-%m-%d"),
         doy = format(monsoon, "%j"),
         monsoonend = as.Date(paste(year, monthend, dayend,
                                    sep = "-"), "%Y-%m-%d"),
         doyend = format(monsoonend,  "%j")) %>% 
  select(year, monsoon, doy, monsoonend, doyend)
head(dhm.monsoon)
# sd.onset <- sd(as.numeric(dhm.monsoon$doy))
# head(dhm.monsoon)
# ggthemr('greyscale')
```

```{r}
p.dhm <- qplot(as.numeric(year),
               as.Date(doy, format="%j",
                       origin="1.1.2013"),
               data=dhm.monsoon,
               # geom=c("xspline"), #method="lm",
               geom=c("line"), #method="lm",
               size=I(1),
               alpha=I(.9),
               main= expression(atop("Official monsoon onset day - Nepal (1968-2020)",
                                     "Data source: DHM media release")),
               xlab="", 
               ylab="Monsoon onset day")
# p.dhm <- p.dhm + stat_smooth(method="lm")
# p.dhm <- p.dhm + geom_line(aes(year, as.Date(doyend, format="%j",
#                        origin="1.1.2013")))
p.dhm <- p.dhm + geom_point(size=I(3))
p.dhm <- p.dhm +
  geom_hline(yintercept = as.numeric(as.Date("162", format="%j",
                       origin="1.1.2013")), size=1,
             colour="firebrick", linetype=2) + 
  scale_x_continuous(breaks=seq(1970, 2020, 5))
p.dhm + theme_hc() +
  theme(legend.position="none")
```

